Learning to Group and Order AMR Semantics for General Purpose Discourse Planning

### 2015-12-10 paper/paper.tex
revisions through preprocessing

\abstract
This paper presents a Machine Learning pipeline for Discourse Planning at a semantic level, looking at sentence and paragraph Abstract Meaning Representations. Due to the flexibility of AMR, the learned planner could be used in a general-purpose NLG system that produces any form of semantically driven text. Our pipeline seeks to learn scoring functions for merging information into higher-level messages, and for ordering the those messages, resulting in a discourse plan. We show that using the learned scoring functions in a discrete-optimization system, we can outperform a random baseline in constructing an ordered sequence of sentence semantics from a semantic world representation, according to an information ordering metric.

@@

\section{Introduction}
Discourse Planners, systems that order and structure the information to be conveyed in a Natural Language Generation system, typically are implemented as hand-written templates, known as `schemas', or by mapping discourse relations to AI planning operators and running an off-the-shelf planner \cite{applied_nlg}. Recent research has shown that Machine Learning techniques can be used to order facts according to a learned scoring function in a restricted domain \cite{learning_to_order_facts}.

@@

An AMR-annotated sentence is directed, rooted, labeled graph, or as a
conjunction of logical triples. Throughout this paper, we work with the AMRs
exclusively in the graphical form $G=(V,E)$, and will thus refer to the AMR
using just the sets $V,E$.

@@

\subsection{Information Ordering}

Information ordering is an integral component of the Discourse Planning phase
of the NLG pipeline, discussed above and in \cite{applied_nlg}. After an
information set has been structured into to pieces of information (``facts" or
sentences), these pieces need to be ordered so that the reader can process and
understand the information being conveyed as easily as possible. This is often
approached as a planning problem, though Dimitromanolaki and
Androutsopoulos\cite{learning_to_order_facts} explore using pipelined decision
trees and instance-based learning.

@@

Early work by Kukich described a system for ``knowledge-based report generation'' in which manually specified domain rules were used to merge database facts into higher level messages, which were then ordered by a discourse module \cite{kukich1983knowledge}. The ordered output could later be passed to generation system to build a report. Work by Paris in 1990 and Hovy in 1993 introduced the idea of using AI Planners to generate discourse plans. They focused on finding explicit discourse relations between messages and using planners to generate an ordering which was allowed under the constraints of the discourse relations identified \cite{paris1990natural}\cite{hovy1993automated}. More recently, Duboue and Mckeown applied Machine Learning (ML) approaches to discourse planning. In \cite{duboue2001empirically}, they present an algorithm to learn ordering constraints among facts. In \cite{duboue2002content}, they build upon this work, using evolutionary algorithms to learn the tree representation of a planner. Although these systems seek to generate valid discourse plans, they do not use a scoring function to choose the `best' among the many valid plans that are available. Dimitromanolaki and Androutsopoulos used ML techniques to learn such a scoring function \cite{learning_to_order_facts}. They train a pipeline of classifiers that repeatedly choose the next fact to be added to a paragraph for a specific domain.
In order to train this pipeline, Dimitromanolaki and Androutsopoulos assume a
fixed $k$ number of sentences per paragraph (in their paper $k=6$).

@@

\section{Data}

The dataset for this research was the AMR ``proxy" dataset - a corpus of
AMR-annotated sentences obtained from newswire data. Since the source text of
the articles was unavailable, we used a sliding window approach to generate
``paragraphs" of size $k$ from the articles. For an article with $n$ sentences,
we could generate paragraphs containing sentences $(s_1,s_2,\ldots,s_k),
(s_2,s_3,\ldots,s_k,s_{k+1}), \ldots (s_{n-k},s_{n-k+1},\ldots,s_n)$.
As an example from our dataset, ``A prosecutor in South Korea seized 80
kilograms of drugs'' would have the following first-order logical form and AMR
representation.\\

\noindent \textbf{Logical form}:
\begin{equation}
\begin{split}
\exists s,d2,p2,c,n: &\\
&\text{instance}(s,want\text{-}01) \wedge \text{instance}(d2, drug) \wedge \\
&\text{instance}(p2, prosecutor) \wedge \text{instance}(c, country) \wedge \\
&\text{instance}(n, name) \wedge \text{arg0}(s,p2) \wedge \text{arg1}(s,d2)
\wedge \\ &\text{arg2}(s,c) \wedge \text{name}(c,n) \wedge \text{op1}(n,
``South") \wedge \\ &\text{op2}(n, ``Korea") \wedge \text{unit}(d2, kilogram)
\wedge \text{quant}(d2, 80)
\end{split}
\end{equation}
\textbf{AMR form}:
\begin{figure}
\includegraphics[width=\linewidth]{amr_example.pdf}
\label{fig:amr_example}
\caption{Example AMR}
\end{figure}

Our training set had 1676 ``paragraphs" of size $k=5$, and our test set had 214
paragraphs.

@@

\subsection{ARG-of Reversal}
The typical annotation for a sentence is the one shown in figure \ref{fig:amr_example}. However, due to the flattened textual representation of AMR, it is sometimes easier for annotators to add reversed edges. For example, in figure \ref{fig:amr_example}, instead of adding the edge: (`s', `p2', `ARG0'), they could have added (`p2', `s', `ARG0-of'), resulting in figure \ref{fig:amr_example_argof}. To fix this, we just reverse any such edge and change the label from `ARGN-of' to `ARGN'.

@@

\subsection{Translation of `and' Semantics}
The next issue that we had to address in the sentence semantics was the way `and' is handled in AMR. For ease of annotation when `and' occurs in a sentence, an `and' instance is created with everything being `anded' together as its children. Then this `and' instance is used as the argument wherever one of the instances children would have gone before. For example, if instead our example sentence were ``A prosecutor in South Korea seized 80 kilograms of drugs and dozens of weapons.'', we would have the AMR graph shown in figure \ref{fig:amr_example_and}.

@@

While this is representation is not necessarily {\em wrong}, the choice to `and' these two entities together was syntactic, not semantic. A strictly semantic representation of this sentence should contain the information: there is a prosecutor, there are drugs, there are weapons, the prosecutor seized the drugs, and the prosecutor seized the weapons. In a generation process, this could be realized as one sentence (``A prosecutor in South Korea seized 80 kilograms of drugs and dozens of weapons.'') or two sentences (``A prosecutor in South Korea seized 80 kilograms of drugs. The prosecutor also seized dozens of weapons.''). Since we ultimately want to look at how information in paragraphs is grouped into sentences and then ordered, we need to translate these `and' semantics back to a strictly semantic representation; otherwise, we have already encoded some of our information grouping.

To recover the actual semantics from the `and' semantics, we must make N copies of each `and' instance, where N is the number of children of that `and' instance. Then, for each child, we must delete the copied `and' instance and replace it with the child. This algorithm is written formally in algorithm \ref{alg:translate_and_semantics}.

When we apply this algorithm to our semantic graph in figure \ref{fig:amr_example_and}, a new `seize-01' instance is made with the `drugs' instance as ARG1 and the `weapons' instance as ARG1 respectively. This resulting graph is shown in figure \ref{fig:amr_example_and_removed}.

@@

\subsection{Graph Merging}
After the semantic graphs have been preprocessed for each sentence, they have to be merged together to form paragraph semantic graphs. The only complication is deciding if two entities in separate sentences are the same. For example if we are merging the sentences ``The red dog chased the cat. Afterwards, the dog slept.'', the two dogs should be recognized as the same entity. However, if the sentences were ``The red dog chased the cat. The black dog slept.'', the two dogs should be recognized as separate.

@@

\[
    equiv(u,v) = \left\{\begin{array}{lr}
    u.\text{label} = v.\text{label}, & u,v \text{ `concepts'} \\
    \forall k \text{ attribute of } u \text{ and } v, v.\text{attr}[k] = v.\text{attr}[k] \text{ } \land, & \text{else} \\
    \forall e,f \text{ childedge of } u,v \text{ with same label, } equiv(e.\text{child}, f.\text{child}) & \\
    \end{array}\right\}
\]

@@

The only issue of concern during graph merging is what to do when a node in one graph has multiple equivalent nodes in the second graph. This can be caused by annotation errors (i.e. a `name' attribute is annotated ``John F. Kennedy'' in one sentence and ``JFK'' in another) or underspecification (i.e. there are two dogs in the world and the next sentence simply refers to ``the dog''). When this happens, there is no clear resolution, so we drop the errant paragraph from our dataset.

>>>>

\abstract
This paper presents a Machine Learning pipeline for Discourse Planning at a semantic level, seeking to generate ordered sentence semantics from unordered paragraph semantics. Due to the flexibility of our semantic representation (Abstract Meaning Representation), the learned planner could be used in a general-purpose NLG system that produces any form of semantically driven text. Our pipeline seeks to learn scoring functions for merging information into higher-level messages, and for ordering those messages which, when applied, result in a discourse plan. We show that using the learned scoring functions in a discrete-optimization system, we can outperform a random baseline in constructing an ordered sequence of sentence semantics from a semantic world representation, according to an information ordering metric.

@@

\section{Introduction}
Discourse Planners, systems that order and structure the information to be conveyed in a Natural Language Generation system, typically are implemented as hand-written templates, known as `schemas', or by mapping discourse relations to AI planning operators and running off-the-shelf planners \cite{applied_nlg}. Recent research has shown that Machine Learning techniques can be used to order facts according to a learned scoring function in a restricted domain \cite{learning_to_order_facts}.

@@

An AMR-annotated sentence is a directed, multi-rooted, labeled graph. This can be represented as a
conjunction of logical triples, of the form (parent\_entity, child\_entity, edge\_label). 
Throughout this paper, we work with the AMRs exclusively in the graphical form $G=(V,E)$, 
and will thus refer to the AMR using just the sets $V,E$.

As an example from our dataset, the sentence ``A prosecutor in South Korea seized 80
kilograms of drugs'' would have the following first-order logical form (Figure \ref{fig:amr_example_logical}) and AMR
representation (Figure \ref{fig:amr_example}).

\begin{figure}
\begin{equation*}
\begin{split}
\exists s,d2,p2,c,n: &\\
&\text{instance}(s,seize\text{-}01) \wedge \text{instance}(d2, drug) \wedge \\
&\text{instance}(p2, prosecutor) \wedge \text{instance}(c, country) \wedge \\
&\text{instance}(n, name) \wedge \text{arg0}(s,p2) \wedge \text{arg1}(s,d2)
\wedge \\ &\text{arg2}(s,c) \wedge \text{name}(c,n) \wedge \text{op1}(n,
``South") \wedge \\ &\text{op2}(n, ``Korea") \wedge \text{unit}(d2, kilogram)
\wedge \text{quant}(d2, 80)
\end{split}
\end{equation*}
\label{fig:amr_example_logical}
\caption{Example AMR in Logical Form}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{amr_example.pdf}
\label{fig:amr_example}
\caption{Example AMR in Graphical Form}
\end{figure}

@@

@@

Early work by Kukich described a system for ``knowledge-based report generation'' in which manually specified domain rules were used to merge database facts into higher level messages, which were then ordered by a discourse module \cite{kukich1983knowledge}. The ordered output could later be passed to generation system to build a report. Work by Paris in 1990 and Hovy in 1993 introduced the idea of using AI Planners to generate discourse plans. They focused on finding explicit discourse relations between messages and using planners to generate an ordering which was allowed under the constraints of the discourse relations identified \cite{paris1990natural}\cite{hovy1993automated}.

More recently, Duboue and Mckeown applied Machine Learning (ML) approaches to discourse planning. In \cite{duboue2001empirically}, they present an algorithm to learn ordering constraints among facts. In \cite{duboue2002content}, they build upon this work, using evolutionary algorithms to learn the tree representation of a planner. Although these systems seek to generate valid discourse plans, they do not use a scoring function to choose the `best' among the many valid plans that are available. Dimitromanolaki and Androutsopoulos used ML techniques to learn such a scoring function \cite{learning_to_order_facts}. They train a pipeline of classifiers, where each classifier chooses the next fact to be added to a domain-specific paragraph. In order to train this pipeline, Dimitromanolaki and Androutsopoulos assumed a fixed $k$ number of sentences per paragraph (in their paper $k=6$).

@@

\section{Data}
\label{sec:data}

The dataset for this research was the AMR ``proxy" dataset - a corpus of
AMR-annotated sentences obtained from newswire data. Since the source text of
the articles was unavailable, we used a sliding window approach to generate
``paragraphs" of size $k$ from the articles. For an article with $n$ sentences,
we could generate paragraphs containing sentences $(s_1,s_2,\ldots,s_k),
(s_2,s_3,\ldots,s_k,s_{k+1}), \ldots (s_{n-k},s_{n-k+1},\ldots,s_n)$.

After applying the sliding window function, our training set had 1676 ``paragraphs" of size $k=5$, and our test set had 214 paragraphs.

@@

\subsection{ARG-of Reversal}
The typical annotation for a sentence is the one shown in Figure \ref{fig:amr_example}. However, due to the annotation format, it is sometimes easier for annotators to add reversed edges. For example, in Figure \ref{fig:amr_example}, instead of adding the edge: (`s', `p2', `ARG0'), an annotator could have added (`p2', `s', `ARG0-of'), resulting in Figure \ref{fig:amr_example_argof}. To fix this, we just reverse any such edge and change the label from `ARGN-of' to `ARGN'.

@@

\subsection{Translation of `and' Semantics}
The next issue that we had to address in the sentence semantics was the way `and' is handled in AMR. For ease of annotation when `and' occurs in a sentence, an `and' instance is created with everything being `anded' together as its children. Then this `and' instance is used as the argument wherever one of the instances children would have been before. For example, if instead our example sentence were ``A prosecutor in South Korea seized 80 kilograms of drugs and dozens of weapons.'', we would have the AMR graph shown in Figure \ref{fig:amr_example_and} instead of the one in Figure \ref{fig:amr_example}.

@@

While this representation is not necessarily {\em wrong}, the choice to `and' these two entities together was syntactic, not semantic. A strictly semantic representation of this sentence should contain the information: there is a prosecutor, there are drugs, there are weapons, the prosecutor seized the drugs, and the prosecutor seized the weapons. In a generation process, this could be realized as one sentence (``A prosecutor in South Korea seized 80 kilograms of drugs and dozens of weapons.'') or two sentences (``A prosecutor in South Korea seized 80 kilograms of drugs. The prosecutor also seized dozens of weapons.''). Since we ultimately want to look at how information in paragraphs is grouped into sentences and then ordered, we need to translate these `and' semantics back to a strictly semantic representation; otherwise, we have already partially encoded our information grouping.

To recover the actual semantics from the `and' semantics, we must make one copy of the subgraph that the `and' instance is an argument to for each child of the `and' instance. Then, for each copy, we delete the `and' instance and replace it with the child. This algorithm is written formally in algorithm \ref{alg:translate_and_semantics}.

When we apply this algorithm to our semantic graph in figure \ref{fig:amr_example_and}, a new `seize-01' instance is made with the `drugs' instance as ARG1 and another with the `weapons' instance as ARG1. This resulting graph is shown in Figure \ref{fig:amr_example_and_removed}.

@@

\subsection{Graph Merging}
After the semantic graphs have been preprocessed for each sentence, they have to be merged together to form a semantic graph representing all of the information communicated by the paragraph. The only complication is deciding if two entities in separate sentences are the same. For example if we are merging the sentences ``The red dog chased the cat. Afterwards, the dog slept.'', the two dogs should be recognized as the same entity. However, if the sentences were ``The red dog chased the cat. The black dog slept.'', the two dogs should be recognized as separate.

@@

\[
    equiv(u,v) = \left\{\begin{array}{lr}
    u.\text{label} = v.\text{label}, & u,v \text{ `concepts'} \\
    \forall k \text{ attribute of } u \text{ and } v, u.\text{attr}[k] = v.\text{attr}[k] \text{ } \land, & \text{else} \\
    \forall e,f \text{ childedge of } u,v \text{ with same label, } equiv(e.\text{child}, f.\text{child}) & \\
    \end{array}\right\}
\]

@@

On potential issue arises when a node in one graph has multiple equivalent nodes in the second graph. This can be caused by annotation errors (i.e. a `name' attribute is annotated ``John F. Kennedy'' in one sentence and ``JFK'' in another) or underspecification (i.e. there are two dogs in the world and the next sentence simply refers to ``the dog''). When this happens, there is no clear resolution, so we drop the errant paragraph from our dataset.
